---
title: "project_first_steps"
output: html_document
---

```{r}
library(tidyverse)
```

STEP 1 - CASES EXTRACTION

```{r}

getwd()
setwd("/rds/general/project/hda_students_data/live/Group7/General/Sarvesh/Scripts")

library(data.table)
library(openxlsx)
library(tidyverse)

# read Biobank dataset
mydata=data.frame(fread("/rds/general/user/ems2817/projects/hda_students_data/live/Group7/General/ukb26390.csv"))
print(mydata$X40006.0.0) # Baseline cancer only, since consequent cancers may be explained by the previous

# determine column of interest
cancer_ICD_col<- (colnames(mydata)[grep("40006",colnames(mydata))])
cancer_ICD_col

# codings dataset load
mycoding=read.csv("/rds/general/user/ems2817/projects/hda_students_data/live/Group7/General/Codings_Showcase.csv")

# look at field of interest (19 = ICD10)
coding_id="19"
print(head(mycoding))
mycoding_field=mycoding[which(mycoding[,1]==coding_id),]
head(mycoding_field)
mycoding_field=mycoding_field[,-1]
rownames(mycoding_field)=mycoding_field[,1]
print(head(rownames(mycoding_field)))

# Recoded categories:
as.character(mycoding_field[as.character(mydata$X40006.0.0),"Meaning"])
mycoding_field[,1]

# Skin cancer
head(mycoding_field)

# Create list of values equated to skin cancer
skincancer =c(c(paste0("C", 430:449)))

# Filter coding fields based on cancer of interest
coding_skincancer = filter(mycoding_field, Value %in% skincancer)
coding_skincancer

# Show results of coding fields that are for cancer of interest
mycoding_skincancer=coding_skincancer[,-1]
rownames(coding_skincancer)=coding_skincancer[,1]
print(coding_skincancer)
coding_skincancer[,1]

codes_skincancer = as.character(coding_skincancer[,1])
print(codes_skincancer)

# Extract skin cancer cases
mydata_C43_C44 = subset(mydata,X40006.0.0 %in% codes_skincancer) 

withdrawn=as.character(read.csv("/rds/general/project/hda_students_data/live/Group7/General/w19266_20200204.csv")[,1])
mydata_C43_C44<-filter(mydata_C43_C44,!(eid %in% withdrawn))


# Save mydata_C43_C44
saveRDS(mydata_C43_C44,"/rds/general/project/hda_students_data/live/Group7/General/Final/mydata_C43_C44.rds")
```

Age of diagnosis of melanoma/skin cancer: 84

Here we need a specific date / age of diagnosis for each participants 
We will compare this to date of comorbidity 

```{r}

Cancer_year <- (colnames(mydata_C43_C44)[grep("84",colnames(mydata_C43_C44))])
print(head(Cancer_year,n=10))

head(mydata_C43_C44$X84.0.0)
head(mydata_C43_C44$X84.0.1)
head(mydata_C43_C44$X84.0.2)
head(mydata_C43_C44$X84.0.3)
head(mydata_C43_C44$X84.0.4)
head(mydata_C43_C44$X84.0.5)

cancer_year_0<-mydata_C43_C44$X84.0.0
```



TO DO --> uniformalize for age 

```{r}

```




```{r}
age_cancer_diagnosis<- (colnames(mydata_C43_C44)[grep("40008",colnames(mydata_C43_C44))])
print(head(age_cancer_diagnosis,n=20))

names(mydata_C43_C44)[names(mydata_C43_C44) == "X40008.0.0"] <- "age_cancer_diagnosis"
sum(is.na(mydata_C43_C44$age_cancer_diagnosis)) # 0 missing values
hist(mydata_C43_C44$age_cancer_diagnosis)
```


STEP 2 - COMORBIDITIES 
 
 A- Non cancer illness-code, self reported
 - Work out how we can extract comorbidities throughout all follow-up times
 
 
 Include all numbers after the second zero
 
```{r}


comorbidity_col<- (colnames(mydata_C43_C44)[grep("20002",colnames(mydata_C43_C44))])
comorbidity_col

codes_comorbidity<-as.factor(mydata_C43_C44$X20002.0.0)

comorbidity_prev<-sort(prop.table(table(codes_comorbidity)), decreasing=TRUE)
top_comorbidity<-data.frame(comorbidity_prev)
colnames(top_comorbidity)<-c('Value','freq')

comorbidity_coding_id="6"
mycoding_field=mycoding[which(mycoding[,1]==comorbidity_coding_id),]
mycoding_field=mycoding_field[,-1]
mycoding_field=data.frame(mycoding_field)


comorbidities_full<- merge(x = top_comorbidity, y = mycoding_field, by = "Value", all = TRUE)
top_comorbidity<-comorbidities_full[1:40,]


```



TO DO: Need to improve the count part to make it more efficient 

```{r}
###extract 20002.0.X
head(mydata)
mydata2 <- mydata
pos_comor <- grep("X20002.0.", fixed=TRUE, colnames(mydata)) #Obtain column position of all comorbidities
comor <- mydata[, pos_comor]

###Function to obtain dictionary for the specific column
mycoding=read.csv("/rds/general/user/ems2817/projects/hda_students_data/live/Group7/General/Codings_Showcase.csv")
rename <- function(x, coding_id){
  mycoding_field=mycoding[which(mycoding[,1]==coding_id),]
  mycoding_field=mycoding_field[,-1]
  mycoding_field$Values <- as.factor(mycoding_field$Value)
  #mycoding_field$Values <- droplevels(mycoding_field$Values)
  mycoding_field <- subset(mycoding_field, mycoding_field[,1] != -1)
  rownames(mycoding_field)=mycoding_field[,1]
  #level_names <- mycoding_field[,2]
  #x <- as.factor(x)
  #levels(x) <- level_names
  return(mycoding_field)
}

###Get comorbidities by real name
table(comor$X20002.0.0)
comor_dict <- rename(comor$X20002.0.0, "6")
#change na to 0 for code to work
mydata2[, pos_comor] [is.na(mydata2[, pos_comor])] <- 0
#substitute coding numbers with disease name
for(i in 1:dim(mycoding_field)[1]){
  mydata2[, pos_comor] <- apply(mydata2[,pos_comor], 2,function(x) {(gsub(comor_dict[i, "Value"], comor_dict[i, "Meaning"], x))})
}

###Function to count frequency each comorbidity (Very inefficient!!!!!!)
all_comorbidity <- mydata2[, pos_comor]
comorbidity_count <- data.frame(mycoding_field[, 2])
for(k in 1:dim(mycoding_field)[1]){
  count <- 0
  for(i in 1:dim(all_comorbidity)[1]){
    for(j in 1:dim(all_comorbidity)[2]){
      if(all_comorbidity[i,j] == mycoding_field[k, 2]){
        count <- count + 1
      } 
    }
  }
  comorbidity_count$count[k] <- count
}


```
 

Date of diagnosis of comorbidity - code 87
```{r}
date_comorbidity="37"
mycoding_date_comorbidity=mycoding[which(mycoding[,1]==date_comorbidity),]

comorbidity_date<- (colnames(mydata_C43_C44)[grep("87",colnames(mydata_C43_C44))])
comorbidity_date

comorbidity_date_0<-mydata_C43_C44$X87.0.0
```

TO DO: uniformalize 


```{r}
diff_outcome_comorbidity <-cancer_year_0-comorbidity_date_0
```

Categorize depending on difference



Comorbidity cancer : 20001 - coding 3

```{r}
cancer_comorbidity="20001"
mycoding_cancer_comorbidity=mycoding[which(mycoding[,1]==cancer_comorbidity),]

cancer_com<- (colnames(mydata_C43_C44)[grep("20002",colnames(mydata_C43_C44))])
cancer_com
```

```{r}
###extract 20002.0.X
head(mydata)
mydata2 <- mydata
pos_comor <- grep("X20004.0.", fixed=TRUE, colnames(mydata)) #Obtain column position of all comorbidities
comor <- mydata[, pos_comor]

###Function to obtain dictionary for the specific column
mycoding=read.csv("/rds/general/user/ems2817/projects/hda_students_data/live/Group7/General/Codings_Showcase.csv")
rename <- function(x, coding_id){
  mycoding_field=mycoding[which(mycoding[,1]==coding_id),]
  mycoding_field=mycoding_field[,-1]
  mycoding_field$Values <- as.factor(mycoding_field$Value)
  #mycoding_field$Values <- droplevels(mycoding_field$Values)
  mycoding_field <- subset(mycoding_field, mycoding_field[,1] != -1)
  rownames(mycoding_field)=mycoding_field[,1]
  #level_names <- mycoding_field[,2]
  #x <- as.factor(x)
  #levels(x) <- level_names
  return(mycoding_field)
}

###Get comorbidities by real name
table(comor$X20002.0.0)
comor_dict <- rename(comor$X20002.0.0, "3")
#change na to 0 for code to work
mydata2[, pos_comor] [is.na(mydata2[, pos_comor])] <- 0
#substitute coding numbers with disease name
for(i in 1:dim(mycoding_field)[1]){
  mydata2[, pos_comor] <- apply(mydata2[,pos_comor], 2,function(x) {(gsub(comor_dict[i, "Value"], comor_dict[i, "Meaning"], x))})
}

###Function to count frequency each comorbidity (Very inefficient!!!!!!)
all_comorbidity <- mydata2[, pos_comor]
comorbidity_count <- data.frame(mycoding_field[, 2])
for(k in 1:dim(mycoding_field)[1]){
  count <- 0
  for(i in 1:dim(all_comorbidity)[1]){
    for(j in 1:dim(all_comorbidity)[2]){
      if(all_comorbidity[i,j] == mycoding_field[k, 2]){
        count <- count + 1
      } 
    }
  }
  comorbidity_count$count[k] <- count
}

```


Cancer year/age of cancer comorbidity 84

```{r}
cancer_comorbidity_year="84"
mycoding_cancer_comorbidity_year=mycoding[which(mycoding[,1]==cancer_comorbidity_year),]

cancer_com_year<- (colnames(mydata_C43_C44)[grep("20002",colnames(mydata_C43_C44))])
cancer_com_year
```


Within people who have comorbidities - did these occur before occurence of cancer?
When we have a comorbidity that occured before melanoma, create categorical variable for time
1
3
5
10


OPTION 2 FOR EXTRACTING COMORBIDITIES

Extract all comorbidity columns
```{r}
mydata <- readRDS("/rds/general/project/hda_students_data/live/Group7/General/Eléonore/cases.rds")
write.csv(mydata, "/rds/general/project/hda_students_data/live/Group7/General/Carolina/cases.csv")
data <- mydata[1:10,]

#load the field id list to be extracted
myfields=unname(unlist(read.table("/rds/general/project/hda_students_data/live/Group7/General/Carolina/comorbidities.txt", header=FALSE)))



# Extracting the column ids 
column_id=grep("eid", colnames(mydata_C43_C44))
found_fieldids=NULL
for (k in 1:length(mydata_C43_C44)){
  mygrep=grep(paste0("X",myfields[k],"."), fixed=TRUE, colnames(mydata_C43_C44))
  if (length(mygrep)>0){
    found_fieldids=c(found_fieldids, myfields[k])
  }
  column_id=c(column_id, mygrep)
}
#remove duplicates
column_id2 <- unique(column_id)

# Extracting required columns from dataset
extracted<-data.frame(mydata_C43_C44) %>%
          select(column_id2) # Path to change!

saveRDS(extracted, "/rds/general/project/hda_students_data/live/Group7/General/Eléonore/extracted_comorb.rds")

#Maintain only at baseline

disease_baseline <- c()
for (k in 1:length(myfields)){
  disease_baseline[k] <-paste0("X",myfields[k],".0.0")
}

extracted_baseline <- extracted$V1
disease_baseline2 <- c()
for(k in 2:length(myfields)){
  for(j in 1:dim(extracted)[2]){
    if (disease_baseline[k] == colnames(extracted)[j]){
      disease_baseline2[k] <- disease_baseline[k]
      extracted_baseline <- cbind(extracted_baseline, mydata_C43_C44[disease_baseline2[k]])
    }
  }
}


disease_count <- c()
column_names <- colnames(extracted_baseline)
for(i in 1:dim(extracted_baseline)[2]){
  disease_count[i] <- sum((extracted_baseline[, i]) >= 1, na.rm = TRUE)
}

disease_count
names(disease_count) <- column_names
sort(disease_count)
sorted_disease_count <- sort(disease_count, decreasing=TRUE)
disease_count_top50 <- sorted_disease_count[1:50]
disease_count_top50_prop <- disease_count_top50/ dim(extracted)[1]
diseases_name <- names(disease_count_top50_prop)
table_disease <- data.frame("disease" = diseases_name, "count" = round(unname(disease_count_top50),3),
                              "prop" = round(unname(disease_count_top50_prop), 3))
table_disease

```

X40006 --> cancer, they all  have melanoma 

Need to recode to data field in order to get comorbidity name 
How do we get coding for each of these?


STEP 3 - COVARIATES EXTRACTION 

AGE - year of birth 

```{r}
DEM_age<- (colnames(mydata_C43_C44)[grep("34",colnames(mydata_C43_C44))])
print(head(DEM_age))

names(mydata_C43_C44)[names(mydata_C43_C44) == "X34.0.0"] <- "DEM_age"


prop.table(table(mydata_C43_C44$DEM_age))

sum(is.na(mydata_C43_C44$DEM_age))
```

ETHNICITY

```{r}
DEM_ethnicity<- (colnames(mydata_C43_C44)[grep("21000",colnames(mydata_C43_C44))])
print(head(DEM_ethnicity))

names(melanoma_cases)[names(mydata_C43_C44) == "X21000.0.0"] <- "DEM_ethnic"


coding_ethnic='1001'
mycoding_field=mycoding[which(mycoding[,1]==coding_ethnic),]
mycoding_field=mycoding_field[,-1]


prop.table(table(melanoma_cases$DEM_ethnic))

melanoma_cases$DEM_ethnic<-as.factor(mycoding_field['Meaning'])

sum(is.na(mydata_C43_C44$DEM_ethnic))
```

SEX

```{r}


DEM_sex<- (colnames(mydata_C43_C44)[grep("31",colnames(mydata_C43_C44))])
print(head(DEM_sex))

names(mydata_C43_C44)[names(mydata_C43_C44) == "X31.0.0"] <- "DEM_sex"

prop.table(table(mydata_C43_C44$DEM_sex))

t=table(mydata_C43_C44$DEM_sex)
a=t[1]
b=t[2]
c=round(a/nrow(mydata_C43_C44)*100 , 2)
d=round(b/nrow(mydata_C43_C44)*100 , 2)

c
d

sum(is.na(mydata_C43_C44$DEM_sex))
```

Describe our population
Look at distribution of individual covariates
Check for missingness
Correlation?

```{r}
#install.packages("ukbtools")

library(ukbtools)
?ukbtools

my_ukb_data<-ukb_df(mydata_C43_C44)
```


STEP 4 - MATCHING - CONTROLS

Remember to remove participants with no consent

```{r}


```


STEP 5-
Data imputation 
research

